---
title       : Reproducible Pitch Presentation
subtitle    : Developing Data Products
author      : Sandipan Dey
job         : Data Scientist
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}

---

* The application is meant to provide a simple visualization for the mtcars dataset.     
  There are two tabs on the web application. 
* The first tab displays some reactive displays of the values of the variable selected from the dropdown. It also displays the summary of the selected variable.
* The number of data tuples displayed can be changed as well.
* The second tab displays linear regression (OLS) model fits for different independent variables 9can be selected from drop down list). Also a 3rd variable   
  value is shown by color coding, which can also be selected from another drop down. Accordingly the OLS model changes.
* The R code snippet on the next slide show the figures that are shown reactively.

--- 


* The first tab provides some basic **exploratory analysis** on a variable, while
  the second tab visualizes a **predictive modeling** (in terms of OLS fitting) to model the variable
  mpg (miles per gallon) as a linear function of the input variable chosen from the drop down.

* If one selects the "mpg" variable on the first tab, the summary and the values for the 
  the variable will be shown in a tabular format as shown below, although the implementation
  will be reactive w.r.t. the variable selected. 

```{r chunkLabel}
summary(mtcars$mpg)
```

---

* If on second tab, one selects the "wt" as the independent variable ans "cyl" as color, the following model will be fit

```{r chunkLabel1}
library(ggplot2)
qplot(wt, mpg, data=mtcars, geom=c("point", "smooth"), 
method="lm", formula=y~x, color=cyl, 
main="Linear Regression (OLS) Fit", 
xlab="wt", ylab="mpg")
```

--- 

* On the other hand, in the second tab, if one selects "qsec" as the independent variable ans "cyl" as color, the following model will be fit.

```{r chunkLabel2}
library(ggplot2)
qplot(qsec, mpg, data=mtcars, geom=c("point", "smooth"), 
method="lm", formula=y~x, color=cyl, 
main="Linear Regression (OLS) Fit", 
xlab="qsec", ylab="mpg")
```


